---
title: "r4openEO: UC2 - Timeseries break detection"
author: 
 - Peter Zellner, Eurac Research
date: "02/11/2022"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: TRUE
---

# Description

Process graph for getting s2 ndvi from Eurac backend, masking the clouds, and running the bfast R-UDF for detecting breaks in the time series. Additionally refining the result with the magnitude of change! Using only one UDF that can be change it's output according to the context that is passed through to it.

-   extent: either test.json (100x100 px) or vaja.json
-   time: 2016 to 2021

# libs

```{r, libs}
library(openeo)
library(dplyr)
library(sf)
library(stars) 
```

# connect to backend

```{r, connect}
host = "https://openeo.eurac.edu" # "https://openeo.cloud" # "https://openeo-dev.vito.be"
con = openeo::connect(host)
login()

# con$isConnected()
# con$isLoggedIn()
# describe_account()
```

# definitions

output file

```{r, name_out}
name_out = "test_magn_mask.tif" 
# name_out = "vaja.tif"
file_format = "GTiff" # "GTiff" # "netCDF"
pth_out = "~/git_projects/r4openeo-usecases/uc2-ts-breakdetection/openeo_eurac/results" 
pth_out = file.path(pth_out, name_out)
pth_out
```

path to spatial extent

```{r, path_ext}
pth_ext = "~/git_projects/r4openeo-usecases/uc2-ts-breakdetection/00_extent/test.json"
# pth_ext = "~/git_projects/r4openeo-usecases/uc2-ts-breakdetection/00_extent/vaja.json"
```

path to udf
```{r, pth_udf}
pth_udf =  "https://raw.githubusercontent.com/Open-EO/r4openeo-usecases/main/uc2-ts-breakdetection/00_udfs/bfast_udf.R"
```


# parameters

```{r, params_pg}
collection = "S2_L2A_ALPS" 
spatial_extent = jsonlite::read_json(pth_ext) 
time_range = list("2016-01-01", "2021-01-01") 
bands = c("B04", "B08", "SCL") 
bands_cloud = "SCL"
```

# define process graph 
```{r, processgraph}
p = openeo::processes()

data = p$load_collection(id = collection, 
                         spatial_extent = list(west = spatial_extent$spatial_extent$west, 
                                               east = spatial_extent$spatial_extent$east, 
                                               south = spatial_extent$spatial_extent$south, 
                                               north = spatial_extent$spatial_extent$north),
                         temporal_extent = time_range, bands = bands)

data_clouds = p$filter_bands(data = data, bands = bands_cloud)

fun_cloud_mask = function(data, context) { 
  scl = data 
  res = scl == 3 | scl == 8 | scl == 9 
  return(res)
  }

cloud_mask = p$apply(data = data_clouds, process = fun_cloud_mask)

mask_clouds = p$mask(data = data, mask = cloud_mask)

calc_ndvi = p$reduce_dimension(data = mask_clouds, 
                               dimension = "bands", 
                               reducer = function(data, context) { 
                                 red = data[1] 
                                 nir = data[2] 
                                 (nir-red)/(nir+red)})

reduce_udf = p$reduce_dimension(data = calc_ndvi, 
                                dimension = "DATE", # at Eurac it's called DATE, at VITO "time" or "t" 
                                reducer = function(data, context){
                                  p$run_udf(data = data, 
                                            context = list(start_monitor = 2018),
                                            udf = pth_udf, 
                                            runtime = "R")})

reduce_udf_magn = p$reduce_dimension(data = calc_ndvi, 
                                dimension = "DATE",
                                reducer = function(data, context){
                                  p$run_udf(data = data, 
                                            context = list(start_monitor = 2018, 
                                                           val = "magnitude"),
                                            udf = pth_udf, 
                                            runtime = "R")})

fun_magn_mask = function(data, context) { 
  magn = data 
  res = magn > -0.2 
  return(res)
  }

magn_mask = p$apply(data = reduce_udf_magn, process = fun_magn_mask)

mask_magn = p$mask(data = reduce_udf, mask = magn_mask)

result = p$save_result(data = mask_magn, format=file_format) # "GTiff" #"netCDF"
```

# look at the pg
```{r, look_at_pg}
pg_json = as(result, "Process")
pg_json
```


# create job

create job

```{r, create_job}
job = create_job(graph = result, 
                 title = name_out, 
                 description = name_out, 
                 format = file_format) 
start_job(job = job$id)
```

status job
```{r, status_job}
status(job) 
```

download
```{r, dwnld}
result_obj = list_results(job = job$id) 
dwnld = download_results(job = job$id, folder = pth_out)
```

# quick check

```{r, quick_check}
file.size(dwnld[[2]])/1000000 # mb 
tst = read_stars(dwnld[[2]]) 
tst[[1]][tst[[1]]<2018] = NA
mapview::mapview(tst)
```

