---
title: "r4opneEO: UC2 - Timeseries break detection"
author: 
 - Peter Zellner, Eurac Research
date: "02/11/2022"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: TRUE
---

# Description

Process graph for getting s2 ndvi from Eurac backend, masking the clouds, and running the bfast R-UDF for detecting breaks in the time series.

-   extent: either test.json (100x100 px) or vaja.json
-   time: 2016 to 2021

# libs

```{r, libs}
library(openeo)
library(dplyr)
library(sf)
library(stars) 
```

# connect to backend

```{r, connect}
host = "https://openeo.eurac.edu" # "https://openeo.cloud" # "https://openeo-dev.vito.be"
con = openeo::connect(host)
login()

# con$isConnected()
# con$isLoggedIn()
# describe_account()
```

# definitions

output file

```{r, name_out}
name_out = "test.tif" 
file_format = "GTiff" # "GTiff" # "netCDF"
pth_out = "~/git_projects/r4openeo-usecases/uc2-ts-breakdetection/openeo_eurac/results" 
pth_out = file.path(pth_out, name_out)
```

path to spatial extent

```{r, path_ext}
pth_ext = "~/git_projects/r4openeo-usecases/uc2-ts-breakdetection/00_extent/test.json"
```

path to udf
```{r, pth_udf}
pth_udf =  "https://raw.githubusercontent.com/Open-EO/r4openeo-usecases/main/uc2-ts-breakdetection/00_udfs/bfast_udf.R"
```


# parameters

```{r, params_pg}
collection = "S2_L2A_ALPS" 
spatial_extent = jsonlite::read_json(pth_ext) 
time_range = list("2016-01-01", "2021-01-01") 
bands = c("B04", "B08", "SCL") 
bands_cloud = "SCL"
```

# define process graph 
```{r, processgraph}
p = openeo::processes()

data = p$load_collection(id = collection, 
                         spatial_extent = list(west = spatial_extent$spatial_extent$west, 
                                               east = spatial_extent$spatial_extent$east, 
                                               south = spatial_extent$spatial_extent$south, 
                                               north = spatial_extent$spatial_extent$north),
                         temporal_extent = time_range, bands = bands)

data_clouds = p$filter_bands(data = data, bands = bands_cloud)

fun_cloud_mask = function(data, context) { 
  scl = data 
  res = scl == 3 | scl == 8 | scl == 9 
  return(res)
  }

cloud_mask = p$apply(data = data_clouds, process = fun_cloud_mask)

mask_clouds = p$mask(data = data, mask = cloud_mask)

calc_ndvi = p$reduce_dimension(data = mask_clouds, 
                               dimension = "bands", 
                               reducer = function(data, context) { 
                                 red = data[1] 
                                 nir = data[2] 
                                 (nir-red)/(nir+red)})

reduce_udf = p$reduce_dimension(data = calc_ndvi, 
                                dimension = "DATE", # at Eurac it's called DATE, at VITO "time" or "t" 
                                reducer = function(data, context){
                                  p$run_udf(data = data, 
                                            context = list(start_monitor = 2018),
                                            udf = pth_udf, 
                                            runtime = "R")})


result = p$save_result(data = reduce_udf, format=file_format) # "GTiff" #"netCDF"
```

# look at the pg
```{r, look_at_pg}
pg_json = as(result, "Process")
```


# create job

create job

```{r, create_job}
job = create_job(graph = result, 
                 title = name_out, 
                 description = name_out, 
                 format = file_format) 
start_job(job = job$id)
```

status job
```{r, status_job}
status(job) 
```

download
```{r, dwnld}
result_obj = list_results(job = job$id) 
dwnld = download_results(job = job$id, folder = pth_out)
```

# quick check

```{r, quick_check}
file.size(dwnld[[1]])/1000000 # mb 
tst = read_stars(dwnld[[1]]) 
tst 
plot(tst)
```

# timings

vaja on vito-dev for 2016 to 2021 (5GB) CPU usage 933,550 cpu-seconds Wall time 4,775 seconds Memory usage 2,377,836,590 mb-seconds Sentinelhub 38,620.25 sentinelhub_processing_unit

test on vito-dev for 2016 to 2021 (10MB) CPU usage 26,126 cpu-seconds Wall time 216 seconds Memory usage 66,473,528 mb-seconds Sentinelhub 386 sentinelhub_processing_unit
